The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2016-02-14 script/de/script-de.tex
First version of CT Book ed 12. Most important point is the newly includes bitciphers chapter. Further things: Tried to reduce latex warnings amap. Changed wording from Sage to SageMath where visible...

% $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,11pt,oneside]{book}
\overfullrule=5pt % uncomment to mark overfull boxes
\usepackage[latin1]{inputenc} % Umlaute
% \usepackage[dvips]{graphics}
% \usepackage{graphicx}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{fancyvrb}
\usepackage{makeidx}
\usepackage[pdftex]{graphicx}
\usepackage{pdfpages}   % SR_2009: added new cover
\usepackage{paralist} %BE_2012: eliminate space after end{itemize} worked
                      %         with compactitem
\usepackage{arydshln} %BE_Jan2013: To use hdashline
\usepackage{verse}    %BE_Feb2013: To use \settowidth{\versewidth}{xxx}

@@

% Formatierung der Fußnoten
\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage[hyphens]{url} % options "hyphens" enables linebreaks in \url{}
% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können
\usepackage{float}
\usepackage{subfig}
\usepackage{shorttoc}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,plainpages=false,hyperfootnotes=false]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels, plainpages=false and, hyperfootnotes=false and place hyperref after float.
%Option debug helps finding problems.

\floatstyle{boxed}
\newfloat{ctsquotefloat}{H}{loq}   % ctsQuotes-from-Books
\floatname{ctsquotefloat}{Zitat}
\newenvironment{ctsquote}{\begin{ctsquotefloat}\center}{\end{ctsquotefloat}}

\floatstyle{ruled}

\newfloat{sagecode}{!ht}{loc}[chapter]
\floatname{sagecode}{Sage-Beispiel}

@@

% \setcounter{tocdepth}{¿Anzahl der Ebenen¿} Default ist 3
% \setcounter{secnumdepth}{¿Nummerierungstiefe¿}
\setcounter{tocdepth}{4}    % Das Inhaltsverzeichnis wird vier Ebenen tief
\setcounter{secnumdepth}{4} % Nummerierung auch bei subsubsection

@@

\usepackage[german]{babel}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\vspace{25pt}
	{\Huge Das CrypTool-Skript: \\
	 \vspace{5pt}
	 Kryptographie, Mathematik und mehr} \\
	\vspace{20pt}
	{\LARGE Hintergrundmaterial und Zusatzinformationen\\
	 zum freien E-Learning-Programm CrypTool\\
	 (mit Code-Beispielen zur Zahlentheorie, geschrieben in Sage)} \\
	\vspace{15pt}
        {\large (11. Auflage -- veröffentlicht mit CrypTool-Version 1.4.31)}
        \vspace{180pt}
}

\author{ %
	{\LARGE (c) Prof.\ Bernhard Esslinger (Mitautor und Herausgeber)}\\[5pt]
	{\LARGE und das CrypTool-Team, 1998-2013}\\[5pt]
	{\LARGE Frankfurt am Main}\\*[15pt]
	\url{www.cryptool.org}
}

@@

% \includepdf[pages=1]{figures/coverv1_de.pdf}
\includepdf[pages=1]{figures/coverv1_de_draft.pdf}

@@

\pagestyle{plain}
\setlength{\fboxrule}{.5mm}
\setlength{\fboxsep}{1.75mm}
\setlength{\footnotesep}{6pt}
\addtolength{\footskip}{8pt}

%space between text and footnote 
\renewcommand\footnoterule{
\vspace{2em}%   <-- one line space between text and footnoterule
%\kern-3\p@
\hrule width .4\columnwidth
\vspace{4pt}
%kern 2.6\p@
}

@@

    Copyright \copyright{} 1998--2013 Bernhard Esslinger and the
    CrypTool Team. Permission is granted to copy,
    distribute and/or modify this document under the terms of the GNU
    Free Documentation License, Version 1.3 or any later version
    published by the Free Software Foundation. A copy of
    the license is included in the section entitled
    \hyperlink{appendix-GNU-fdl}{``GNU Free Documentation License''}.

    Dies beinhaltet auch den Code für die Sage-Beispiele in diesem Dokument.
\end{quote}

\vspace{420pt}
Quelle Coverfoto: \url{www.photocase.com}, Andre Günther\\

Schriftsatz-Software: \LaTeX

Versionsverwaltungs-Software: Subversion

@@

%\parindent 0cm
% \parskip 4pt
% Setting the default for the vertical distance before a paragraph as distinguished variable
\newcounter{mycounterDefaultParskip}
\setcounter{mycounterDefaultParskip}{4}
\parskip \value{mycounterDefaultParskip} pt

\input{chapters/introduction.tex}

\mainmatter

\input{chapters/cryptomethods.tex}
\input{chapters/paper_and_pencil.tex}
\input{chapters/primes.tex}
\input{chapters/numbertheory.tex}
\input{chapters/moderncryptography.tex}
\input{chapters/digitalsignatures.tex}
\input{chapters/ellipticcurves.tex}
\input{chapters/homomorphicencryption.tex}
\bibliographyunit[\chapter]
\bibliography*{references}
\bibliographystyle*{plain}
\input{chapters/crypto2020.tex}
\bibliographyunit

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begin{appendix}
\newpage
\hypertarget{appendix-start}{}\label{s:appendix-start}
\chapter{Anhang}
    \begin{itemize}
      \item[1] \hyperlink{appendix-menu-overview-CT1}{CrypTool-1-Menübaum}
      \item[2] \hyperlink{appendix-template-overview-CT2}{CrypTool-2-Vorlagen}
      \item[3] \hyperlink{appendix-function-overview-JCT}{JCrypTool-Funktionen}
      \item[4] \hyperlink{appendix-function-overview-CTO}{CrypTool-Online-Funktionen}
      \item[5] \hyperlink{appendix-movies}{Filme und belletristische
                          Literatur mit Bezug zur Kryptographie}
      \item[6] \hyperlink{appendix-Learn-NT}
                         {Lernprogramm Elementare Zahlentheorie}
      \item[7] \hyperlink{appendix-using-sage}
                         {Kurzeinführung in das Computer-Algebra-System Sage}
      \item[8] \hyperlink{appendix-authors}{Autoren des CrypTool-Skripts}
    \end{itemize}
  \input{chapters/menus.tex}
  \input{chapters/movies.tex}
  \input{chapters/what-is-sage.tex}
  \input{chapters/authors.tex}
\end{appendix}

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begingroup
  % remove extra vertical space from \listoffigure and \listoftables so that 
  % it look similar to \listof{sagecode}
  \setlength{\parskip}{0pt}

@@

  \clearpage\phantomsection
  \addcontentsline{toc}{chapter}{Verzeichnis der Krypto-Verfahren mit Pseudocode}
  \listof{cryptoprocedure}{Verzeichnis der Krypto-Verfahren mit Pseudocode}

\endgroup

@@

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Verzeichnis der Sage-Programmbeispiele}
\listof{sagecode}{Verzeichnis der Sage-Programmbeispiele}
\index{Sage!Programmbeispiele}
\label{sc:List-of-Sage-Code-Examples}
% \hypertarget{sc:List-of-Sage-Code-Examples}{}

\vspace{30pt}
\begin{itemize}
  \item Der Quellcode der Sage-Beispiele in diesem Skript wird in Form von
        Sage-Programmdatei"-en mit dem CrypTool 1-Setup-Programm ausgeliefert.
        Die Beispiele in einem Kapitel sind jeweils in einer Datei zusammen gefasst.\\
        Nach der Installation von CrypTool 1 \index{CrypTool 1} finden Sie
        die Sage-Beispiele im Unterverzeichnis \verb#sage# in den folgenden
        4 Dateien:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
  \item Alle Beispiel wurden mit der Sage-Version 5.3 (Release Date 2012-09-08) getestet.
\end{itemize}   % xxxxxxxxxxxxxx Weitere Beispiele ergänzen !

>>>>

% $Id$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% Add lengthy comments only into the German version! %%%%%%%%%%

\documentclass[a4paper,11pt,oneside]{book}
\overfullrule=5pt % uncomment to mark overfull boxes
\usepackage[latin1]{inputenc} % Umlaute  %%BERM2: Added for biciphers.tex (necessary any more?)
\usepackage[german]{babel} % TODO: ngerman (für neue dt. Rechtschreibung) führte zu Compiliationsfehler
%\usepackage[T1]{fontenc} %BERM2: biciphers.tex used T1.
                          % Ohne T1 wurden dort \dh und \textperthousand nicht erkannt.
                          % Die Hinzunahme von T1 (soll eigentlich Umbrüche verbessern),
                          % erzeugt aber ein paar zusätzliche schwarze Balken (z.B. bei
                          % "Der DES-Algorithmus ist" in cryptomethods.text im Deutschen,
                          % und bei der Trennung von "straddling checker-board" in 2.2.1 im Englischen.
% Statt T1 als Package zu inkludieren, nur die benötigten Sonderzeichen definiert.
\DeclareTextSymbolDefault{\dh}{T1}
\DeclareTextSymbolDefault{\textperthousand}{T1}

\usepackage{datetime} % To distinguish often compiled draft versions
% \usepackage[dvips]{graphics}
% \usepackage{graphicx}

\usepackage{amsmath}
\usepackage{amssymb,amscd,latexsym}   %BERM2: added for biciphers.tex
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{color}   %BERM2: added for biciphers.tex

\usepackage{fancyvrb}
\usepackage{makeidx}
\usepackage[pdftex]{graphicx}
\usepackage{pdfpages}   % SR_2009: added new cover
%\usepackage{enumitem} %BE_2012: eliminate space after end{itemize} didn't work
\usepackage{paralist}  %BE_2012: eliminate space after end{itemize} worked with compactitem
\usepackage{arydshln} %BE_Jan2013: To use hdashline
\usepackage{verse}    %BE_Feb2013: To use \settowidth{\versewidth}{xxx}

@@

% Formatierung der Fußnoten
\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage[hyphens]{url} % options "hyphens" enables linebreaks nach Bindestrich in \url{}
\appto\UrlBreaks{\do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j
\do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t\do\u\do\v\do\w
\do\x\do\y\do\z}  % Trennung in langen URLs nach jedem Kleinbuchstaben möglich.

% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können

\usepackage{float}
\usepackage{subfig}
\usepackage{shorttoc}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,plainpages=false,hyperfootnotes=false]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels, plainpages=false and, hyperfootnotes=false and place hyperref after float.
%Option debug helps finding problems.

\floatstyle{boxed}  % This is necessary to have borders around the Zitate.
                    % After the ctsquote def overruled by \floatstyle{ruled} ?
\newfloat{ctsquotefloat}{H}{loq}   % ctsQuotes-from-Books, List-of-Quotes
\floatname{ctsquotefloat}{Zitat}   % ctsQuotes with centering
\newenvironment{ctsquote}{\begin{ctsquotefloat}\center}{\end{ctsquotefloat}}
\floatstyle{ruled}

\newfloat{sagecode}{!ht}{loc}[chapter]
\floatname{sagecode}{SageMath-Beispiel}

@@

% \newenvironment{ Latex_Create-Closed-Dummy-Environ }{beg09asas} {endasas}
% \begin{ Latex_Create-Closed-Dummy-Environ }
%	TR (this line can be omitted!)
% \end{ Latex_Create-Closed-Dummy-Environ }
% OUTPUT: beg09asas TR endasas
% The commands which don't effect anything outside the environment have to be between begin and end.
% -- analog zu
% \newenvironment{ dummerspruch }[1]{#1 sagt: \begin{quote}}{\end{quote}}
% \begin{ dummerspruch }[ Gollum ]
% My precioussss !
% \end{ dummerspruch }
\newenvironment{ Latex_Create-Closed-Dummy-Environ }{}{}

\setcounter{tocdepth}{4}    % Das Inhaltsverzeichnis wird vier Ebenen tief (Default ist 3)
\setcounter{secnumdepth}{4} % Nummerierung auch bei subsubsection

@@

% Define title ......................
% \title{ % Keine Leerzeilen, sonst Meldung: "Paragraph ended before \title was complete."
% \title{\vspace{25pt}
\title{
% be_Wished to enlarge only the main title, however \HUGE not available,
%    or \HUGE without any effect beside adding \usepackage[11pt]{moresize}
%    and explicit use of \fontsize caused the LaTeX Font Warning:
%    Font shape `OT1/cmr/m/n' in size <40> not available size <24.88> substituted on input line 130.
%    {\usefont{\encodingdefault}{\rmdefault}{b}{n}%
%     \fontsize{40}{48} \selectfont Das CrypTool-Buch}
%    {\fontsize{40}{48} \selectfont Das CrypTool-Buch} \\[15pt]
%    {\HUGE Das CrypTool-Buch} \\[15pt]
% --> Später: Use the KOMA-script class or the memoir class (erfordert aber andere \documentclass)
%             http://texblog.org/2012/08/29/changing-the-font-size-in-latex/
	{\Huge Das CrypTool-Buch \\[15pt]
	       Kryptographie, Mathematik und mehr} \\[25pt]
	{\LARGE Hintergrundmaterial und Zusatzinformationen\\
	 zum freien E-Learning Krypto-Programm CrypTool\\[5pt]
	 (mit Code-Beispielen, geschrieben in SageMath)} \\[25pt]
	%{\large 12. Auflage}
	{\large 12. Auflage -- Draft-Version (\settimeformat{hhmmsstime}\currenttime)}
        \vspace{180pt}
}

\author{ % Die Klammer zur Fontänderung funktioniert bei author nur einzeilig im Gegensatz zu title?!
	{\LARGE Prof.\ Bernhard Esslinger (Mitautor und Herausgeber)}\\[10pt]
	{\LARGE und das CrypTool-Team, 1998-2016}\\[10pt]
	{\LARGE Frankfurt am Main}\\*[15pt]
	\url{www.cryptool.org}
}

@@

% \includepdf[pages=1]{figures/coverv1_de-300dpi.pdf}
% \includepdf[pages=1]{figures/coverv1_de.pdf}
\includepdf[pages=1]{figures/coverv1_de_draft.pdf}

@@

\pagestyle{plain}
\setlength{\fboxrule}{.5mm}
\setlength{\fboxsep}{1.75mm}
\setlength{\footnotesep}{6pt}
\addtolength{\footskip}{8pt}
%\setlength{\footskip}{4cm}

%space between text and footnote 
%\renewcommand{\footnoterule}{\parindent0cm\rule{13cm}{.1pt}\vspace{.2cm}}
%Formatierung der Fu"snoten
\renewcommand\footnoterule{
\vspace{2em}%   <-- one line space between text and footnoterule
%\kern-3\p@
\hrule width .4\columnwidth
\vspace{4pt}
%kern 2.6\p@
}

@@

    Copyright \copyright{} 1998--2016 Bernhard Esslinger and the
    CrypTool Team. Permission is granted to copy,
    distribute and/or modify this document under the terms of the GNU
    Free Documentation License, Version 1.3 or any later version
    published by the Free Software Foundation. A copy of
    the license is included in the section entitled
    \hyperlink{appendix-GNU-fdl}{``GNU Free Documentation License''}.

    Dies beinhaltet auch den Code für die SageMath-Beispiele in diesem Dokument.
\end{quote}

\vspace{420pt}
\noindent
Quelle Coverfoto: \url{www.photocase.com}, Andre Günther\\
Schriftsatz-Software: \LaTeX\\
Versionsverwaltungs-Software: Subversion

@@

% \parindent 0cm
% \parskip 4pt
% Setting the default for the vertical distance before a paragraph as distinguished variable
\newcounter{mycounterDefaultParskip}
\setcounter{mycounterDefaultParskip}{4}
\parskip \value{mycounterDefaultParskip} pt

\input{chapters/introduction.tex}
\mainmatter
\input{chapters/cryptomethods.tex}
\input{chapters/paper_and_pencil.tex}
\input{chapters/primes.tex}
\input{chapters/numbertheory.tex}
\input{chapters/moderncryptography.tex}
\input{chapters/digitalsignatures.tex}
\input{chapters/ellipticcurves.tex}
\input{chapters/bitciphers.tex}
\input{chapters/homomorphicencryption.tex}
\bibliographyunit[\chapter]
\bibliography*{references}
\bibliographystyle*{plain}
\input{chapters/crypto2020.tex}
\bibliographyunit

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begin{appendix}
\newpage
\hypertarget{appendix-start}{}\label{s:appendix-start}
\chapter{Anhang}
    \begin{itemize}
      \item[1] \hyperlink{appendix-menu-overview-CT1}{CrypTool-1-Menübaum}
      \item[2] \hyperlink{appendix-template-overview-CT2}{CrypTool-2-Vorlagen}
      \item[3] \hyperlink{appendix-function-overview-JCT}{JCrypTool-Funktionen}
      \item[4] \hyperlink{appendix-function-overview-CTO}{CrypTool-Online-Funktionen}
      \item[5] \hyperlink{appendix-movies}{Filme und belletristische
                          Literatur mit Bezug zur Kryptographie}
      \item[6] \hyperlink{appendix-Learn-NT}
                         {Lernprogramm Elementare Zahlentheorie}
      \item[7] \hyperlink{appendix-using-sage}
                         {Kurzeinführung in das Computer-Algebra-System SageMath}
      \item[8] \hyperlink{appendix-authors}{Autoren des CrypTool-Skripts}
    \end{itemize}
  \input{chapters/menus.tex}
  \input{chapters/movies.tex}
  \input{chapters/what-is-sage.tex}
  \input{chapters/authors.tex}
\end{appendix}

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\begingroup
  % remove extra vertical space from \listoffigure and \listoftables so that 
  % it looks similar to \listof{sagecode}
  \setlength{\parskip}{0pt}

@@

  \clearpage\phantomsection
  \addcontentsline{toc}{chapter}{Verzeichnis der Krypto-Verfahren mit Pseudocode}
  \listof{cryptoprocedure}{Verzeichnis der Krypto-Verfahren mit Pseudocode}
\endgroup

@@

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Verzeichnis der SageMath-Programmbeispiele}
\listof{sagecode}{Verzeichnis der SageMath-Programmbeispiele}
\index{SageMath!Programmbeispiele}
\label{sc:List-of-Sage-Code-Examples}
% \hypertarget{sc:List-of-Sage-Code-Examples}{}

\vspace{30pt}
\begin{itemize}
  \item Der Quellcode der SageMath-Beispiele in diesem Skript wird in Form von
        SageMath-Programmdatei"-en mit dem CrypTool 1-Setup-Programm ausgeliefert.
        Die Beispiele in einem Kapitel sind jeweils in einer Datei zusammen gefasst.\\
        Nach der Installation von CrypTool 1 \index{CrypTool 1} finden Sie
        die SageMath-Beispiele im Unterverzeichnis \verb#sagemath# in den folgenden
        4 Dateien:\\
        - SageMath-Samples-in-Chap01.sage \\
        - SageMath-Samples-in-Chap02.sage \\
        - SageMath-Samples-in-Chap03.sage \\
        - SageMath-Samples-in-Chap04.sage
  \item Alle Beispiel wurden mit der SageMath-Version 5.3 (Release Date 2012-09-08) xxxxxxxxxxxxxxxxxxxx  getestet.
\end{itemize}   % xxxxxxxxxxxxxx Weitere Beispiele ergänzen !

### 2016-02-14 script/en/chapters/aboutcryptool.tex
First version of CT Book ed 12. Most important point is the newly includes bitciphers chapter. Further things: Tried to reduce latex warnings amap. Changed wording from Sage to SageMath where visible...

% $Id$
% ............................................................................
%                 TEXT OF THE 2nd PAGE
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

In this {\em script accompanying the CrypTool programs} \index{CrypTool}
you will find predominantly mathematically oriented information on using
cryptographic procedures. Also included are many sample code pieces written in the
computer algebra system {\bf Sage}\index{Sage} (see appendix~\ref{s:appendix-using-sage}).
The main chapters have been written by various {\bf authors}
(see appendix~\ref{s:appendix-authors}) %\hyperlink{appendix-authors}{authors}
and are therefore independent from one another. At the end of most chapters
you will find references and web links.
The sections have been enriched with many {\em footnotes}. Within the footnotes
you can see where the described functions can be called in the different CrypTool
versions.

@@

%A \hyperlink{Chapter_Hashes-and-Digital-Signatures}{further chapter}
Chapter \ref{Chapter_Hashes-and-Digital-Signatures} gives
an overview about the status of attacks against modern {\bf hash algorithms}
and is then shortly devoted to {\bf digital signatures}, 
which are an essential component of e-business applications.

@@

>>>>

% $Id$
% ............................................................................
%                 TEXT OF THE 2nd PAGE (Overview)
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

In this {\em script accompanying the CrypTool programs} \index{CrypTool}
you will find predominantly mathematically oriented information on using
cryptographic procedures. Also included are many sample code pieces written in the
computer algebra system {\bf SageMath}\index{SageMath} (see appendix~\ref{s:appendix-using-sage}).
The main chapters have been written by various {\bf authors}
(see appendix~\ref{s:appendix-authors}) %\hyperlink{appendix-authors}{authors}
and are therefore independent from one another. At the end of most chapters
you will find references and web links.
The sections have been enriched with many {\em footnotes}. Within the footnotes
you can see where the described functions can be called in the different CrypTool
versions.

@@

Chapter \ref{Chapter_Hashes-and-Digital-Signatures} gives
an overview about the status of attacks against modern {\bf hash algorithms}
and is then shortly devoted to {\bf digital signatures}, 
which are an essential component of e-business applications.

@@

Chapter \ref{Chapter_BitCiphers} introduces {\bf Boolean algebra}.
This is the foundation for most modern symmetric encryption algorithms
as these operate on bit streams and bit groups.

### 2016-02-14 script/en/chapters/authors.tex
First version of CT Book ed 12. Most important point is the newly includes bitciphers chapter. Further things: Tried to reduce latex warnings amap. Changed wording from Sage to SageMath where visible...

\item[Bernhard Esslinger,] \mbox{}\\
Initiator of the CrypTool project, main author of this script, head Cryptography Competence Center / Infrastructure IT Security Solutions at Deutsche Bank, and professor for IT security and cryptography at the University of Siegen.\\
E-mail: bernhard.esslinger@gmail.com, esslinger@fb5.uni-siegen.de.

@@

@@

\item[Minh Van Nguyen,] \mbox{}\\
Sage developer and documentation quality reviewer.

>>>>

\item[Bernhard Esslinger,] \mbox{}\\
Initiator of the CrypTool project, main author of this script. Professor for IT security and cryptography at the University of Siegen. Former CISO of SAG AG and head IT security at Deutsche Bank.\\
E-mail: bernhard.esslinger@gmail.com, bernhard.esslinger@uni-siegen.de

@@

\item[Klaus Pommerening,] \mbox{}\\
Author of chapter~\ref{Chapter_BitCiphers} (``\nameref{Chapter_BitCiphers}''), 
Professor for mathematics and computer science at Johannes-Gutenberg-Universit\"at. Retired.

@@

\item[Minh Van Nguyen,] \mbox{}\\
SageMath developer and documentation quality reviewer.

### 2016-02-14 script/en/chapters/cryptomethods.tex
First version of CT Book ed 12. Most important point is the newly includes bitciphers chapter. Further things: Tried to reduce latex warnings amap. Changed wording from Sage to SageMath where visible...

% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
\newpage
\hypertarget{CM_Appendix_SageCode}{}
\section{Appendix: Examples using Sage}
\label{CM_Sage_samples}
\index{Sage!Code examples}

\noindent Below is Sage source code related to contents of the
chapter~\ref{Label_Kapitel_1} (``\nameref{Label_Kapitel_1}'').

Further details concerning cryptosystems within Sage (e.g. S-DES) can be
found e.g. in the thesis of Minh Van Nguyen \cite{cm:Nguyen2009}.

@@

The Sage module \texttt{crypto/block\_cipher/miniaes.py} supports Mini-AES to allow
students to explore the inner working of a modern block cipher.

@@

How to use Mini-AES is exhaustively described at the Sage reference page:\\
\url{http://www.sagemath.org/doc/reference/sage/crypto/block_cipher/miniaes.html}.

The following Sage code~\ref{cryptomethods:Mini-AES:Sage_example}
is taken from the release tour of Sage 4.1%
\footnote{
See \url{http://mvngu.wordpress.com/2009/07/12/sage-4-1-released/}
}
and calls the implementation of the Mini-AES.

@@

% ---------------------------------------------------------------------------
\subsection{Further symmetric crypto algorithms in Sage}
\label{CM_Sage_SymCryptoAlg}

The Sage Reference v5.4.1%
\footnote{
See \url{http://www.sagemath.org/doc/reference/sage/crypto/}
}
lists the following further ciphers:\\ 
- Linear feedback shift register (LFSR),\\
- ShrinkingGeneratorCryptosystem,\\
- Blum-Blum-Shub (BBS) pseudorandom bit generator,\\
- Lattice related functions.

@@

\bibitem[Nguyen2009]{cm:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       See Sage publications \url{http://www.sagemath.org/library-publications.html},\\
       \url{http://www.sagemath.org/files/thesis/nguyen-thesis-2009.pdf},\\
       \url{http://sites.google.com/site/nguyenminh2/honours-thesis-2009.pdf}

>>>>

% ---------------------------------------------------------------------------
% ---------------------------------------------------------------------------
\newpage
\hypertarget{CM_Appendix_SageCode}{}
\section{Appendix: Examples using SageMath}
\label{CM_Sage_samples}
\index{SageMath!Code examples}

\noindent Below is SageMath source code related to contents of the
chapter~\ref{Label_Kapitel_1} (``\nameref{Label_Kapitel_1}'').

Further details concerning cryptosystems within SageMath (e.g. S-DES) can be
found e.g. in the thesis of Minh Van Nguyen \cite{cm:Nguyen2009}.

@@

The SageMath module \texttt{crypto/block\_cipher/miniaes.py} supports Mini-AES to allow
students to explore the inner working of a modern block cipher.

@@

How to use Mini-AES is exhaustively described at the SageMath reference page:\\
\url{http://www.sagemath.org/doc/reference/sage/crypto/block_cipher/miniaes.html}.

The following SageMath code~\ref{cryptomethods:Mini-AES:Sage_example}
is taken from the release tour of SageMath 4.1%
\footnote{
See \url{http://mvngu.wordpress.com/2009/07/12/sage-4-1-released/}
}
and calls the implementation of the Mini-AES.

@@

% ---------------------------------------------------------------------------
\subsection{Further symmetric crypto algorithms in SageMath}
\label{CM_Sage_SymCryptoAlg}

The SageMath Reference v5.4.1%
\footnote{
See \url{http://www.sagemath.org/doc/reference/sage/crypto/}
}
lists the following further ciphers:\\ 
- Linear feedback shift register (LFSR),\\
- ShrinkingGeneratorCryptosystem,\\
- Blum-Blum-Shub (BBS) pseudorandom bit generator,\\
- Lattice related functions.

@@

\bibitem[Nguyen2009]{cm:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       See SageMath publications \url{http://www.sagemath.org/library-publications.html},\\
       \url{http://www.sagemath.org/files/thesis/nguyen-thesis-2009.pdf},\\
       \url{http://sites.google.com/site/nguyenminh2/honours-thesis-2009.pdf}
